The Effect of Authoritarian Legislature on Regime-Opposition interaction

### 2014-12-03 writeups/argument_overview.tex
preparing slides

This article argues that if the authoritarian legislature indeed helps institutionalize the regime-dissident interaction, then we should see decreased level of violence between the two sides in the presence of an authoritarian legislature. Taking a page from the rationalist war literature, I explains how political violence should only happen when the regime and the opposition misjudge one another's capacity and preference, or when the bargaining game does not have the option of side payments. Given transparent information, a commitment device, and regular channels to negotiate side payments, reaching an agreement is always preferable to violence due to the savings in the social cost of violence.

Armed with this theoretical argument, I empirically examine the level of violence in regime-dissident interaction using a novel event dataset across authoritarian regimes post Cold War. I construct the dataset by looking explicitly at countries that create a legislature during their authoritarian spells and matched these countries with those of the same conditions but do not create legislature. In addition to this large-n investigation, I zoom in at the causal mechanism with a survey experiment implemented in Vietnam, testing whether the recent inclusion of businessmen as delegates in the National Assembly prompts business to participate in the legislature. This survey experiment achieves two goals. First, it allows us to have a cleaner causal inference compared to large-n observational data, especially since violence is potentially too complex a process to model completely. Second, it investigates the immediate behavioral response of business to co-optation attempts, which serves as the logical pre-requisite of any secondary material outcomes that relies on the regime and the dissident interacting with each other. I find that the presence of a legislature does reduce violence across regimes, albeit with a modest substantive effect size since the model's flexibility lets most of the effects is captured in country and country-year intercepts. On the other hand, the survey experiment conclusively shows that the legislature successfully works as a co-optation device, prompting more willingness to participate in the legislature from business. These two findings together complete the missing links in our understanding of authoritarian institutions. They ascertain that the legislature does foster positive regime-opposition interaction, a necessary condition for regime durability, increased investment, and economic growth.

>>>>

This article argues that if the authoritarian legislature indeed helps institutionalize the regime-dissident interaction, then we should see decreased level of violence between the two sides in the presence of an authoritarian legislature. Taking a page from the rationalist war literature, I explain how political violence should only happen when the regime and the opposition misjudge each other's capacity and preference, or when the bargaining game does not have the option of side payments. Given transparent information, a commitment device, and regular channels to negotiate side payments, reaching an agreement is always preferable to violence due to the savings in the social cost of violence.

Armed with this theoretical argument, I empirically examine the level of violence in regime-dissident interaction using a novel event dataset across authoritarian regimes post Cold War. I construct the dataset by looking explicitly at countries that create a legislature during their authoritarian spells and matched these countries with those with the same conditions but do not create legislature. In addition to this large-n investigation, I zoom in at the causal mechanism with a survey experiment implemented in Vietnam, testing whether the recent inclusion of businessmen as delegates in the National Assembly prompts business to participate in the legislature. This survey experiment achieves two goals. First, it allows us to have a cleaner causal inference compared to large-n observational data, especially since violence is potentially too complex a process to model completely. Second, it investigates the immediate behavioral response of business to co-optation attempts, which serves as the logical pre-requisite to any secondary material outcomes that relies on the regime and the dissident interacting with each other. I find that the presence of a legislature does reduce violence across regimes, albeit with a modest substantive effect size since the model's flexibility lets most of the effects be captured in country and country-year intercepts. On the other hand, the survey experiment conclusively shows that the legislature successfully works as a co-optation device, prompting more willingness to participate in the legislature from business. These two findings together complete the missing links in our understanding of authoritarian institutions. They ascertain that the legislature does foster positive regime-opposition interaction, a necessary condition for regime durability, increased investment, and economic growth.

### 2014-12-03 writeups/introduction.tex
preparing slides

While our understanding of authoritarian institution has branched to numerous substantive topics, its causal root deserves more careful studies. Given its common definition as the ``rule of the game,'' institutions are fundamentally a set of constraints on human interactions, and thus its effect must be most evident in the interaction between strategic actors itself \citep{North1990}. And yet most studies have focused on the secondary socioeconomic effects of authoritarian legislature instead of investigating the dynamics of regime-opposition interaction, which is causally prior \citep{Gandhi2007, Svolik2012}. The goal of this article is, therefore, to empirically examine this interaction, arguing that authoritarian legislature should reduce poliltical violence by providing information, lengthening the bargaining game, and enabling side payments.

>>>>

While our understanding of authoritarian institution has branched to numerous substantive topics, its causal root deserves more careful studies. Given the common definition as the ``rule of the game,'' institutions are fundamentally a set of constraints on human interactions, and thus its effect must be most evident in the interaction between strategic actors itself \citep{North1990}. And yet most studies have focused on the secondary socioeconomic effects of authoritarian legislature instead of investigating the dynamics of regime-opposition interaction, which is causally prior \citep{Gandhi2007, Svolik2012}. The goal of this article is, therefore, to empirically examine this interaction, arguing that authoritarian legislature should reduce poliltical violence by providing information, lengthening the bargaining game, and enabling side payments.

### 2014-12-03 writeups/multilevel_datadescription.tex
preparing slides

As my theoretical argument emphasizes the role of the legislature as an informational and commitment device, a country-year is coded as as having a legislature if multiple parties are legal. Even if opposition parties do not win seat, the electoral process also conveys information about the organizational strength and the preference of both sides. This coding rule operationalizes as having LIEC (Legislative Index of Electoral Competitiveness) be 5 or higher in the DPI (Database of Political Institutions).\footnote{The values LIEC are as follows: 4---1 party, multiple candidates; 5---multiple parties are legal but only one party won seats; 6---multiple parties did win seats but the largest party received more than 75\% of the seats; 7---largest party got less than 75\% \citep[14]{Keefer2002}.}

@@

Using a dataset of machine coded news reports around the world from 1991-2012, I am able to look at individual events that happened between the regime and the dissident, enabling an analysis at a highly granular level. These event data come from ICEWS (Integrated Crisis Early Warning System), a project funded by DARPA (Defense Advanced Research Projects Agency) to aid the US policy makers in predicting and responding to crises around the world . These news articles are harvested from ``over 75 international sources (AP, UPI, and BBC Monitor) as well as regional sources (India Today, Jakarta Post, Pakistan Newswire, and Saigon Times),'' ensuring comprehensive coverage \citep[94]{OBrien2010, OBrien2013}. While this dataset has been de-duplicated once so that an event covered by multiple news report only shows up once, I created my own filter to delete duplicates that share the same source actor, target actor, and happen at the same time and place.

>>>>

As my theoretical argument emphasizes the role of the legislature as an informational and commitment device, a country-year is coded as as having a legislature if multiple parties are legal. Even if opposition parties do not win seat, the electoral process still conveys information about the organizational strength and the preference of both sides. This coding rule operationalizes as having LIEC (Legislative Index of Electoral Competitiveness) be 5 or higher in the DPI (Database of Political Institutions).\footnote{The values LIEC are as follows: 4---1 party, multiple candidates; 5---multiple parties are legal but only one party won seats; 6---multiple parties did win seats but the largest party received more than 75\% of the seats; 7---largest party got less than 75\% \citep[14]{Keefer2002}.}

@@

Using a dataset of machine coded news reports around the world from 1991-2012, I am able to look at individual events that happened between the regime and the dissident, enabling an analysis at a highly granular level. These event data come from ICEWS (Integrated Crisis Early Warning System), a project funded by DARPA (Defense Advanced Research Projects Agency) to aid the US policy makers in predicting and responding to crises around the world. These news articles are harvested from ``over 75 international sources (AP, UPI, and BBC Monitor) as well as regional sources (India Today, Jakarta Post, Pakistan Newswire, and Saigon Times),'' ensuring comprehensive coverage \citep[94]{OBrien2010, OBrien2013}. While this dataset has been de-duplicated once so that an event covered by multiple news report only shows up once, I created my own filter to delete duplicates that share the same source actor, target actor, and happen at the same time and place.

### 2014-12-03 writeups/multilevel_empiricalstrategy.tex
preparing slides

Second, it is important to recognize that an authoritarian regime without a legislature is not the correct counter factual, and thus make a very poor control group. Indeed, a regime may not create a co-opting legislature because there is never a sufficient threat of violence from the opposition. Therefore, even though we may see that regimes without legislature are very peaceful, this does not mean that a legislature has no effect on regime-dissident interaction. Conversely, a regime under rising violent threat may institute a legislature that successfully puts a lid on the growing discontent. However, since the level of violence is now constant, it will look as though the legislature had no effect.

>>>>

Second, it is important to recognize that an authoritarian regime without a legislature is not the correct counter factual, and thus makes a very poor control group. Indeed, a regime may not create a co-opting legislature because there is never a sufficient threat of violence from the opposition. Therefore, even though we may see that regimes without legislature are very peaceful, this does not mean that a legislature has no effect on regime-dissident interaction. Conversely, a regime under rising violent threat may institute a legislature that successfully puts a lid on the growing discontent. However, since the level of violence is now constant, it will look as though the legislature had no effect.

### 2014-12-03 writeups/theory.tex
preparing slides

To consider whether violent means is chosen to reach their goal, we need a theoretical framework to understand how the opposition and the regime engage in the bargaining game. A possible starting point is the bargaining models developed for democratic legislatures, where bargaining outcomes are driven by various institutional rules (positive and negative rule, sequencing of these powers, first-mover advantage, etc.) \citep{Baron1989, Cox2006}. However, this type model is not suited to analyze the bargaining under an authoritarian regime, where the institutional rules themselves are under constant threat of being revised by the final arbiter that is violence. In other words, the end game of legislative bargaining now has an extra node: whether either side will violently suppress the other’s demand. Given this context, the decision of the autocrat and the opposition is better characterized as a bargaining game in anarchy, very much like the bargaining between two sovereign states with the option of going to war \citep{Fearon1995}.

To formally model the interaction, let the negotiation between the opposition and the regime be about the share $x$, $(0 < x < 1)$ of a good (normalized as 1) given to the opposition.\footnote{For concreteness we can think of this as the share of the economy’s rent or capital stock} Both sides have the option 1) to bargain and decide on $x$, or 2) to engage in violent protest / repression. Violence is costly compared to peaceful negotiation---we call this difference in relative cost $c_A$ and $c_O$ for the autocrat and the opposition, respectively. If violence happens, whichever side emerges victorious decides the share $x$ to his heart’s desire---i.e. the winning autocrat would give 0 to the opposition and the winning opposition would give 1 to itself.

Given the uncertain nature of collective action battles, the utility of engaging in violence for the two sides is best formulated as an expected utility that depends on the probability of the opposition winning (call this probability $p$). Thus, the expected utility of the opposition is $p – c_O$, while the expected utility of the ruling regime is $1 - p - c_A$. These two expected utilities sum to less than one ($1 - c_A - c_O < 1$), which raises the puzzle why two sides fail to engage in ex-ante bargaining that splits the difference, giving both sides a higher payoff than the expected utility of violence. Indeed, any amount of x between $(p - c_O, p + c_A)$, i.e. the bargaining range, is more desirable than violence to both sides.

@@

\subsection{Legislature increasing violence}
While the prevalent theory of legislature as an informational and commitment device favors hypothesis H1, it is important to note that there are reasons to susepct the opposite. Throughout the previous model, it is implied that the actors’ utility functions remain static with the only parameters being the cost of violence and the share of the issue space. Yet this assumption may miss the dynamism of the opposition’s democratic expectation, which becomes increasingly demanding as the regime starts liberalizing.

>>>>

To consider whether violent means is chosen to reach their goal, we need a theoretical framework to understand how the opposition and the regime engage in the bargaining game. A possible starting point is the bargaining models developed for democratic legislatures, where bargaining outcomes are driven by various institutional rules (positive and negative rule, sequencing of these powers, first-mover advantage, etc.) \citep{Baron1989, Cox2006}. However, this type model is not suited to analyze the bargaining under an authoritarian regime, where the institutional rules themselves are under constant threat of being revised by the final arbiter that is violence. In other words, the end game of legislative bargaining now has an extra node: whether either side will violently suppress the other's demand. Given this context, the decision of the autocrat and the opposition is better characterized as a bargaining game in anarchy, very much like the bargaining between two sovereign states with the option of going to war \citep{Fearon1995}.

To formally model the interaction, let the negotiation between the opposition and the regime be about the share $x$, $(0 < x < 1)$ of a good (normalized as 1) given to the opposition.\footnote{For concreteness we can think of this as the share of the economy’s rent or capital stock} Both sides have the option 1) to bargain and decide on $x$, or 2) to engage in violent protest / repression. Violence is costly compared to peaceful negotiation---we call this difference in relative cost $c_A$ and $c_O$ for the autocrat and the opposition, respectively. If violence happens, whichever side emerges victorious decides the share $x$ to his heart’s desire---e.g. the winning autocrat would give 0 to the opposition and the winning opposition would give 1 to itself.

Given the uncertain nature of collective action battles, the utility of engaging in violence for the two sides is best formulated as an expected utility that depends on the probability of the opposition winning (call this probability $p$). Thus, the expected utility of the opposition is $p - c_O$, while the expected utility of the ruling regime is $1 - p - c_A$. These two expected utilities sum to less than one ($1 - c_A - c_O < 1$), which raises the puzzle why two sides fail to engage in ex-ante bargaining that splits the difference, giving both sides a higher payoff than the expected utility of violence. Indeed, any amount of x between $(p - c_O, p + c_A)$, i.e. the bargaining range, is more desirable than violence to both sides.

@@

\subsection{Legislature increasing violence}
While the prevalent theory of legislature as an informational and commitment device favors hypothesis H1, it is important to note that there are reasons to suspect the opposite. Throughout the previous model, it is implied that the actors' utility functions remain static with the only parameters being the cost of violence and the share of the issue space. Yet this assumption may miss the dynamism of the opposition's democratic expectation, which becomes increasingly demanding as the regime starts liberalizing.
